Relational Parametricity for Algebraically Indexed Types

### 2012-05-29 paper/section-general-framework.tex
Make the notation for inference rules with varying arity more consistent
Finish the description of simultaneous substitutions

We now define simultaneous substitutions of index terms. Given a pair
of index contexts $\Delta$ and $\Delta' = i_1 \mathord: s_1, ..., i_n
\mathord: s_n$, a simultaneous substitution $\sigma : \Delta
\Rightarrow \Delta'$ is a sequence of terms $\sigma = (e_1,...,e_n)$
such that for each $j$ such that $1 \leq j \leq n$, $\Delta \vdash e_j
: s_j$. Given a simultaneous substitution $\sigma = (e_1,...,e_n) :
\Delta \Rightarrow \Delta'$ and a particular variable $i_j : s_j$ in
$\Delta'$, we write $\sigma(i_j)$ to stand for the index term $e_j$.

\begin{enumerate}
\item Define the action of simultaneous substitutions on index terms
  and types: $\sigma^*e$ and $\sigma^*A$. State that simultaneous
  substitution preserves well-sorted-ness and well-indexed-ness.
\item Define composition of simultaneous substitutions
\item Define lifting of simultaneous substitutions
\item Define projection simultaneous substitutions
\end{enumerate}

\begin{lemma}
  Let $\sigma : \Delta \Rightarrow \Delta'$ be a simultaneous
  substitution.
  \begin{enumerate}
  \item If $\Delta' \vdash e : s$, then $\Delta \vdash \sigma^*e : s$; and
  \item If $\Delta' \vdash A \isType$, then $\Delta \vdash \sigma^*A
    \isType$.
  \end{enumerate}
\end{lemma}

@@

  \inferrule*
  {\Delta \vdash e_1 \equiv e'_1 : s_1 \\
    \dots \\
    \Delta \vdash e_n \equiv e'_n : s_n}
  {\Delta \vdash \indexOp{f}(e_1, ..., e_n) \equiv \indexOp{f}(e'_1, ..., e'_n) : s}
\end{mathpar}
We also assume the standard reflexivity, symmetry and transitivity
rules for equality.

@@

>>>>

We now define simultaneous substitutions of index terms. Given a pair
of index contexts $\Delta$ and $\Delta' = i_1 \mathord: s_1, ..., i_n
\mathord: s_n$, a simultaneous substitution $\sigma : \Delta
\Rightarrow \Delta'$ is a sequence of terms $\sigma = (e_1,...,e_n)$
such that for each $j$ such that $1 \leq j \leq n$, $\Delta \vdash e_j
: s_j$. Given a simultaneous substitution $\sigma = (e_1,...,e_n) :
\Delta \Rightarrow \Delta'$ and a variable $i_j \mathord: s_j$ in
$\Delta'$, we write $\sigma(i_j)$ for the index term $e_j$.

For a context $\Delta = i_1\mathord:s_1,...,i_n\mathord:s_n$, and a
variable/type pair $i\mathord:s$ such that $i$ does not appear in
$\Delta$, we define the \emph{projection} simultaneous substitution
$\pi_{i\mathord:s} : \Delta,i\mathord:s \Rightarrow \Delta$ as
$\pi_{i\mathord:s} = (i_1,...,i_n)$. Note that the subscript on
$\pi_{i\mathord:s}$ is the variable/type pair that is being projected
away.

For a simultaneous substitution $\sigma : \Delta \Rightarrow \Delta'$
where $\Delta' = i_1\mathord:s_1,...,i_n\mathord:s_n$ and a
variable/type pair $i\mathord:s$ such that $i$ does not appear in
either $\Delta$ or $\Delta'$, then we can form the \emph{lifted}
simultaneous substitution $\sigma_{i\mathord:s} = (\sigma(i_1), ...,
\sigma(i_n), i) : \Delta,i\mathord:s \Rightarrow
\Delta',i\mathord:s$. Note that we have implicity used the fact that
well-sortedness of index terms is preserved by weakening.

Given a simultaneous substitution $\sigma : \Delta \Rightarrow
\Delta'$ and a well-sorted index term $\Delta' \vdash e : s$, we can
apply $\sigma$ to $e$ to produce a new index term $\sigma^*e$. The new
index term $\sigma^*e$ is defined by the following two clauses:
\begin{mathpar}
  \sigma^*i \isDefinedAs \sigma(i)

  \sigma^*(\indexOp{f}(e_1,...,e_n)) \isDefinedAs \indexOp{f}(\sigma^*e_1, ..., \sigma^*e_n)
\end{mathpar}
Similarly, given a well-indexed type $\Delta' \vdash A \isType$, we
can apply $\sigma$ to $A$ to produce a new type $\sigma^*A$. The
application of a simultaneous substitution to a type is defined
recursively by the following clauses:
\begin{mathpar}
  \sigma^*(\tyPrim{X}{e_1,...,e_n}) \isDefinedAs \tyPrim{X}{\sigma^*e_1,...,\sigma^*e_n}

  \sigma^*\tyUnit \isDefinedAs \tyUnit

  \sigma^*(A \tyArr B) \isDefinedAs \sigma^*A \tyArr \sigma^*B

  \sigma^*(A \tyProduct B) \isDefinedAs \sigma^*A \tyProduct \sigma^*B

  \sigma^*(A + B) \isDefinedAs \sigma^*A + \sigma^*B

  \sigma^*(\forall i\mathord:s.A) \isDefinedAs \forall i\mathord:s.\sigma_{i\mathord:s}^*A
\end{mathpar}
\begin{lemma}
  Let $\sigma : \Delta \Rightarrow \Delta'$ be a simultaneous
  substitution.
  \begin{enumerate}
  \item If $\Delta' \vdash e : s$, then $\Delta \vdash \sigma^*e : s$; and
  \item If $\Delta' \vdash A \isType$, then $\Delta \vdash \sigma^*A
    \isType$.
  \end{enumerate}
\end{lemma}

The \emph{composition} of two simultaneous substitutions $\sigma :
\Delta \Rightarrow \Delta'$ and $\sigma' : \Delta' \Rightarrow \Delta$
is defined as $(\sigma^*e'_1, ..., \sigma^*e'_n)$, where $\sigma' =
(e'_1,...,e'_n)$.

@@

  \inferrule*
  {\{\Delta \vdash e_j \equiv e'_j : s_j\}_{1\leq j\leq n}}
  {\Delta \vdash \indexOp{f}(e_1, ..., e_n) \equiv \indexOp{f}(e'_1, ..., e'_n) : s}
\end{mathpar}
We also assume the standard reflexivity, symmetry and transitivity
rules for equality.

@@

We define a pair of simultaneous substitutions $\sigma, \sigma' :
\Delta \Rightarrow \Delta'$ to be judgmentally equal if their
component terms are judgmentally equal in the context $\Delta$: i.e.~
if $\Delta \vdash e_j \equiv e'_j : s_j$, for all $j$. We write
$\Delta \vdash \sigma \equiv \sigma' : \Delta'$ when two simultaneous
substitutions are judgmentally equal.
